---
title: Hierarchical Conformal Prediction with an application to MERFISH mouse ileum
  data
author: "Daniela Corbetta"
date: "2024-06-27"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Code to reproduce analysis in section 4 of the paper "Conformal inference for cell type annotation with graph-structured constraints"

# Load libraries and utility

```{r, message=FALSE}
library(SingleCellExperiment)
library(scater)
library(scran)
library(VGAM)
library(igraph)
library(ontoProc)
library(MerfishData)
library(ExperimentHub)
library(kableExtra)
library(scConform)
library(BiocParallel)
library(tidyr)
library(ggplot2)
`%notin%` = Negate(`%in%`)
```


# Load data and build ontology

Load data and build ontology (for later)

```{r}
# Load data
spe_baysor <- MouseIleumPetukhov2021(
    segmentation = "baysor",
    use.images = FALSE, use.polygons = FALSE
)
# Load ontology
cl <- getOnto("cellOnto", "2023")

tags <- c(
    "CL:0009022", # Stromal
    "CL:0000236", # B cell
    "CL:0009080", # Tuft
    "CL:1000411", # Endothelial
    "CL:1000335", # Enterocyte
    "CL:1000326", # Goblet
    "CL:0002088", # ICC
    "CL:0009007", # Macrophage + DC
    "CL:1000343", # Paneth
    "CL:0000669", # Pericyte
    "CL:1000278", # Smooth Muscle
    "CL:0009017", # Stem + TA
    "CL:0000492", # T (CD4+)
    "CL:0000625", # T (CD8+)
    "CL:0017004" # Telocyte
)
opi <- graph_from_graphnel(onto_plot2(cl, tags))

## Delete instances from other ontologies
sel_ver <- V(opi)$name[c(grep("CARO", V(opi)$name), grep("BFO", V(opi)$name))]
opi1 <- opi - sel_ver

## Rename vertex to match annotations
V(opi1)$name[grep("CL:0000236", V(opi1)$name)] <- "B cell"
V(opi1)$name[grep("CL:1000411", V(opi1)$name)] <- "Endothelial"
V(opi1)$name[grep("CL:1000335", V(opi1)$name)] <- "Enterocyte"
V(opi1)$name[grep("CL:1000326", V(opi1)$name)] <- "Goblet"
V(opi1)$name[grep("CL:0002088", V(opi1)$name)] <- "ICC"
V(opi1)$name[grep("CL:0009007", V(opi1)$name)] <- "Macrophage + DC"
V(opi1)$name[grep("CL:1000343", V(opi1)$name)] <- "Paneth"
V(opi1)$name[grep("CL:0000669", V(opi1)$name)] <- "Pericyte"
V(opi1)$name[grep("CL:1000278", V(opi1)$name)] <- "Smooth Muscle"
V(opi1)$name[grep("CL:0009017", V(opi1)$name)] <- "Stem + TA"
V(opi1)$name[grep("CL:0009022", V(opi1)$name)] <- "Stromal"
V(opi1)$name[grep("CL:0000492", V(opi1)$name)] <- "T (CD4+)"
V(opi1)$name[grep("CL:0000625", V(opi1)$name)] <- "T (CD8+)"
V(opi1)$name[grep("CL:0017004", V(opi1)$name)] <- "Telocyte"
V(opi1)$name[grep("CL:0009080", V(opi1)$name)] <- "Tuft"

## Add the edge from connective tissue cell and telocyte and delete redundant
## nodes
opi1 <- add_edges(opi1, c("connective\ntissue cell\nCL:0002320", "Telocyte"))
gr <- as_graphnel(opi1)
opi2 <- opi1 - c(
    "somatic\ncell\nCL:0002371", "contractile\ncell\nCL:0000183",
    "native\ncell\nCL:0000003"
)

V(opi2)$name <- trimws(gsub("CL:.*|\\n", " ", V(opi2)$name))

gr1 <- as_graphnel(opi2)

## Plot the final ontology
attrs <- list(node = list(shape = "box", fontsize = 15, fixedsize = FALSE))
plot(gr1, attrs = attrs)
```

Preprocess data:
```{r}
spe_baysor$cell_type <- spe_baysor$leiden_final
spe_baysor$cell_type[spe_baysor$cell_type %in% c(
    "B (Follicular, Circulating)",
    "B (Plasma)"
)] <- "B cell"
spe_baysor$cell_type[grep("Enterocyte", spe_baysor$cell_type)] <- "Enterocyte"
spe_baysor <- spe_baysor[, spe_baysor$cell_type %notin% c(
    "Removed",
    "Myenteric Plexus"
)]
spe_baysor

# See frequencies of cell types
table(spe_baysor$cell_type)
```
Add quality control metrics:
```{r}
spe_baysor <- addPerCellQCMetrics(spe_baysor)
```


## Fit a classification model
- Randomly select 500 cells for train dataset
- Fit logit multinomial model on the train dataset
- Split the remaining cells into calibration (1000 cells) and test. See performance of the model on the test dataset

```{r model, warning=FALSE}
# Randomly select 500 cells
set.seed(1703)
train <- sample(1:ncol(spe_baysor), 500)

# Training data
spe_train <- spe_baysor[, train]
spe_train
spe_other <- spe_baysor[, -train]

# get HVGs
spe_train <- logNormCounts(spe_train)
v <- modelGeneVar(spe_train)
hvg <- getTopHVGs(v, n = 50)

# Extract counts and convert data into a data.frame format
df_train <- as.data.frame(t(as.matrix(counts(spe_train[hvg, ]))))
df_other <- as.data.frame(t(as.matrix(counts(spe_other[hvg, ]))))
df_train$Y <- spe_train$cell_type
df_other$Y <- spe_other$cell_type
table(df_train$Y)

# Fit multinomial model
fit <- vglm(Y ~ .,
    family = multinomial(refLevel = "B cell"),
    data = df_train
)

# split data
set.seed(1237)
cal <- sample(1:nrow(df_other), 1000)
df_cal <- df_other[cal,]
df_test <- df_other[-cal,]

# See performance on the test data
p_test <- predict(fit, newdata=df_test, type="response")
pr_class <- apply(p_test, 1, function(row) colnames(p_test)[which.max(row)])
mean(pr_class==df_test$Y)
```

# Compare standard conformal inference with conformal risk control

Obtain prediction sets with standard conformal inference and the graph-based method.

```{r}
p_cal <- predict(fit, newdata=df_cal, type="response")
# Conformal inference
sets_cf <- getPredictionSets(p_test, p_cal, df_cal$Y, follow_ontology = FALSE,
                             resample = FALSE, return_sc = FALSE, 
                             labels = unique(df_cal$Y))
# Check coverage
(cvg_cf <- mean(mapply(function(vec, lst) vec %in% lst, df_test$Y, sets_cf)))

# Graph-based
sets_graphs <- getPredictionSets(p_test, p_cal, df_cal$Y, follow_ontology = TRUE,
                             resample = FALSE, return_sc = FALSE, onto = opi2,
                             labels = unique(df_cal$Y),
                             BPPARAM = MulticoreParam(workers = 4))
# Check coverage
(cvg_graphs <- mean(mapply(function(vec, lst) vec %in% lst, df_test$Y, sets_graphs)))
```


## Comparison of sizes of prediction sets

```{r}
sizes_std <- sapply(sets_cf, length)
sizes_gr <- sapply(sets_graphs, length)
sizes_std_df <- as.data.frame(table(sizes_std))
sizes_gr_df <- as.data.frame(table(sizes_gr))
names(sizes_std_df) <- c("size", "Split")
names(sizes_gr_df) <- c("size", "Graphs")
merged <- merge(sizes_std_df, sizes_gr_df, by='size', all.x = TRUE, all.y=TRUE)

merged$Graphs[is.na(merged$Graphs)] <- 0
merged$Split[is.na(merged$Split)] <- 0

table(sizes_std)
table(sizes_gr)

mean(sizes_std)
mean(sizes_gr)

long_df <- pivot_longer(merged, cols = c("Split", "Graphs"), 
                        names_to = "Group", values_to = "Count")
long_df$size <- as.factor(long_df$size)  # for categorical x-axis
long_df$Group <- factor(long_df$Group, levels = c("Split", "Graphs"))

# Fig.3 in the paper
ggplot(long_df, aes(x = size, y = Count, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  scale_fill_manual(values = c("Split" = "#B2DF8A", "Graphs" = "#CAB2D6")) +
  labs(x = "Set Size", y = "Frequency", fill = "Method") +
  theme_minimal(base_size = 14)
```
Bigger prediction sets with graphs

## Comparison of homogeneity of elements inside of prediction sets

```{r}
# Obtain distance matrix
Dshortest <- distances(opi2)
avg.dist.intra.gr <- avg.dist.intra.gr2 <- avg.dist.intra.std <- rep(NA, length(sets_cf))
for(i in 1:length(sets_graphs)){
  dist.gr <- as.matrix(Dshortest[sets_graphs[[i]], sets_graphs[[i]]])
  dist.std <- as.matrix(Dshortest[sets_cf[[i]], sets_cf[[i]]])
  avg.dist.intra.gr[i] <- sum(dist.gr)/(ncol(dist.gr)*(ncol(dist.gr)-1))
  avg.dist.intra.std[i] <- sum(dist.std)/(ncol(dist.std)*(ncol(dist.std)-1))
}

avg.dist.intra.gr[is.na(avg.dist.intra.gr)] <- 0
avg.dist.intra.std[is.na(avg.dist.intra.std)] <- 0

mean(avg.dist.intra.gr, na.rm=T)
mean(avg.dist.intra.std, na.rm=T)

# Lower when working with graphs
```

```{r}
s_h_gr <- data.frame(distance=avg.dist.intra.gr, size=sizes_gr)
mean(s_h_gr$distance[s_h_gr$size==1])
boxplot(s_h_gr$distance~s_h_gr$size)

s_h_std <- data.frame(distance=avg.dist.intra.std, size=sizes_std)
mean(s_h_std$distance[s_h_std$size==1])
boxplot(s_h_std$distance~s_h_std$size)

avg.dist.intra.gr
sizes_gr
```


```{r}
library(ggplot2)
library(ggdist)
# Combine data
s_h_gr$method <- "Graphs"
s_h_std$method <- "Split"
combined <- rbind(s_h_gr, s_h_std)

# Convert size to factor if not already
combined$size <- factor(combined$size)

combined <- combined[combined$size!=1,]
combined$method <- factor(combined$method, levels = c("Split", "Graphs"))
# Plot


ggplot(combined, aes(x = size, y = distance)) +
  geom_boxplot(aes(fill = method), alpha=0.5) +
  geom_swarm(
    aes(color = method, group = method), 
    position = position_dodge(width = 0.75)
  ) +
  labs(title = "Intra-group Distance by Size and Method",
       x = "Size", y = "Distance") +
  scale_fill_manual(values = c("Graphs" = "#CAB2D6", "Split" = "#B2DF8A")) +
  scale_color_manual(values = c("Graphs" = "#6A3D9A", "Split" = "#33A02C")) + 
  theme_minimal()
```








```{r}
unique(sets_graphs[sizes_gr==7])
avg.dist.intra.gr[sizes_gr==4]
```


# Calibration set size

Consideration on the appropriate size of the calibration set. In the
standard case, there is a result that states
$$ P(Y_{test} \in C(X_{test})| \{(X_i, Y_i)_{i=1}^{n_{cal}}\}) \sim Beta(n+1-l,l),$$
with $l=\lfloor(n+1)\alpha\rfloor$. In our case there is not a similar result.
To evaluate the appropriate calibration set size, apply the algorithm with $n_{cal}\in\{10,50,100,500,1000\}$
on $B=100$ different calibration/test splits

```{r}
conf.pred.n <- function(B, n){
  cvg.crc <- cvg.conf <- rep(NA, B)
  for(i in 1:B){
    cal <- sample(1:nrow(df_other), n)
    df.cal <- df_other[cal,]
    df.test <- df_other[-cal,]
    p.cal <- predict(fit, newdata=df.cal, type="response")
    p.test <- predict(fit, newdata=df.test, type="response")
    sets_cf <- getPredictionSets(p.test, p.cal, df.cal$Y, follow_ontology = FALSE,
                                 resample = FALSE, return_sc = FALSE, 
                                 labels = unique(df_cal$Y))
    cvg.conf[i] <- mean(mapply(function(vec, lst) vec %in% lst, df.test$Y, sets_cf))
    
    sets_graphs <- getPredictionSets(p.test, p.cal, df.cal$Y, follow_ontology = TRUE,
                                     resample = FALSE, return_sc = FALSE, onto = opi2,
                                     labels = unique(df_cal$Y),
                                     BPPARAM = MulticoreParam(workers = 6))
    cvg.crc[i] <- mean(mapply(function(vec, lst) vec %in% lst, df.test$Y, sets_graphs))
    if(i%%10==0) cat(i)
  }
  return(list(cvg.conf=cvg.conf, cvg.crc=cvg.crc))
}
```

```{r}
set.seed(1221)
n100 <- conf.pred.n(B=100, n=100)
#save(n100, file="n100.RData")
n500 <- conf.pred.n(B=100, n=500)
#save(n500, file="n500.RData")
n1000 <- conf.pred.n(B=100, n=1000)
#save(n1000, file="n1000.RData")
```

```{r}
lapply(n100, mean)
lapply(n500, mean)
lapply(n1000, mean)

lapply(n100, sd)
lapply(n500, sd)
lapply(n1000, sd)

# Theoretical dist
var.beta <- function(alpha=0.1, n){
  l <- floor((n+1)*alpha)
  ((n+1-l)*l)/((n+1)^2*(n+2))
}

mean.beta <- function(alpha=0.1, n){
  l <- floor((n+1)*alpha)
  (n+1-l)/(n+1)
}

n_list <- c(100, 500, 1000)

th.var <- sapply(n_list, function(n) var.beta(n=n))
round(th.var,5)
th.mean <- sapply(n_list, function(n) mean.beta(n=n))
th.mean

emp.var <- c(
  var(n100[[1]]),
  var(n500[[1]]),
  var(n1000[[1]])
)

emp.var.crc <- c(
  var(n100[[2]]),
  var(n500[[2]]),
  var(n1000[[2]])
)

emp.m <- c(
  mean(n100[[1]]),
  mean(n500[[1]]),
  mean(n1000[[1]])
)

emp.m.crc <- c(
  mean(n100[[2]]),
  mean(n500[[2]]),
  mean(n1000[[2]])
)


```

```{r}
plot(1, type = "n",
     xlim = c(min(n_list)-1, max(n_list)+1), 
     ylim = c(
       min(c(emp.m - sqrt(emp.var),
             emp.m.crc - sqrt(emp.var.crc),
             th.mean - sqrt(th.var))),
       max(c(emp.m + sqrt(emp.var),
             emp.m.crc + sqrt(emp.var.crc),
             th.mean + sqrt(th.var)))
     ),
     xlab = "Calibration set size",
     ylab = "Values",
     main = "Means ± Standard Deviations"
)

points(n_list, emp.m, pch = 16, col = "red")
for (i in seq_along(n_list)) {
  arrows(n_list[i], emp.m[i] - sqrt(emp.var)[i],
         n_list[i], emp.m[i] + sqrt(emp.var)[i],
         angle = 90, code = 3, length = 0.1, col = "red")
}

points(n_list, emp.m.crc, pch = 17, col = "blue")
for (i in seq_along(n_list)) {
  arrows(n_list[i], emp.m.crc[i] - sqrt(emp.var.crc)[i],
         n_list[i], emp.m.crc[i] + sqrt(emp.var.crc)[i],
         angle = 90, code = 3, length = 0.1, col = "blue")
}

points(n_list, th.mean, pch = 15, col = "darkgreen")
for (i in seq_along(n_list)) {
  arrows(n_list[i], th.mean[i] - sqrt(th.var)[i],
         n_list[i], th.mean[i] + sqrt(th.var)[i],
         angle = 90, code = 3, length = 0.1, col = "darkgreen")
}

abline(h = 0.9, lty = 2)
legend("topleft",
       legend = c("empirical", "CRC", "theoretical"),
       col = c("red", "blue", "darkgreen"),
       pch = c(16, 17, 15),
       bty = "n")

```

```{r}
library(ggplot2)

# Larger staggering since n_list = c(100,500,1000)
offset_emp <- -20
offset_crc <-   0
offset_th  <- +20

df_plot <- data.frame(
  n = rep(n_list, 3),
  mean = c(emp.m,          emp.m.crc,          th.mean),
  sd   = c(sqrt(emp.var),  sqrt(emp.var.crc),  sqrt(th.var)),
  type = rep(c("emp_mean", "crc_mean", "th_mean"), each = length(n_list)),
  sd_type = rep(c("emp_sd", "crc_sd", "th_sd"), each = length(n_list)),
  x = c(n_list + offset_emp,
        n_list + offset_crc,
        n_list + offset_th)
)

p.meansd <- ggplot() +
  # Means
  geom_point(data = df_plot, 
             aes(x = x, y = mean, color = type),
             size = 3) +
  
  # SD bars with their own colors
  geom_errorbar(data = df_plot,
                aes(x = x, ymin = mean - sd, ymax = mean + sd, color = sd_type),
                width = 10, alpha = 0.9) +
  
  geom_hline(yintercept = 0.9, linetype = "dashed") +
  
  scale_color_manual(
    values = c(
      # Means
      "emp_mean" = "#33A02C",
      "crc_mean" = "#6A3D9A",
      "th_mean"  = "#7F7F7F",
      # SD bars
      "emp_sd" = "#B2DF8A",
      "crc_sd" = "#CAB2D6",
      "th_sd"  = "#CFCFCF"
    ),
    breaks = c("emp_mean","crc_mean","th_mean"),
    labels = c("Split","Graphs","Theoretical")
  ) +
  
  scale_x_continuous(
    breaks = n_list,
    labels = n_list
  ) +
  
  labs(
    x = "Calibration set size",
    y = "Values",
    color = ""
  ) +
  
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )+ggtitle("Empirical mean coverage with standard errors")

n_values <- c(10,50,100,500,1000)
alpha <- 0.1
ccol1 <- c("red", "yellow", "green", "blue", "purple")
plot(0, type='n', xlim=c(0.8, 1), ylim=c(0, 44), xlab='x', ylab='Density', 
     main="Beta density")

for (i in 1:length(n_values)) {
  n <- n_values[i]
  a <- n + 1 - floor((n + 1) * alpha)
  b <- floor((n + 1) * alpha)
  curve(dbeta(x, a, b), add = TRUE, col = ccol1[i], n=10^6)
}
abline(v=1-alpha, lty=2, col="grey")
legend("topright", legend = n_values, col = ccol1, lty = 1, title="n")

```


```{r}
n <- 100
alpha <- 0.1
a <- n + 1 - floor((n + 1) * alpha)
b <- floor((n + 1) * alpha)
curve(dbeta(x, a, b), col = "#7F7F7F", n=10^6, xlim=c(0.82,0.98), ylim=c(0,18),
      main="n_cal=100", xlab="", ylab="")
lines(density(n100$cvg.conf), col = "#33A02C")
lines(density(n100$cvg.crc), col="#6A3D9A")

n <- 500
alpha <- 0.1
a <- n + 1 - floor((n + 1) * alpha)
b <- floor((n + 1) * alpha)
curve(dbeta(x, a, b), col = "#7F7F7F", n=10^6, xlim=c(0.82,0.98),
       main="n_cal=500", xlab="", ylab="")
lines(density(n500$cvg.conf), col = "#33A02C")
lines(density(n500$cvg.crc), col="#6A3D9A")

n <- 1000
alpha <- 0.1
a <- n + 1 - floor((n + 1) * alpha)
b <- floor((n + 1) * alpha)
curve(dbeta(x, a, b), col = "#7F7F7F", n=10^6, xlim=c(0.82,0.98),
       main="n_cal=1000", xlab="", ylab="")
lines(density(n1000$cvg.conf), col = "#33A02C")
lines(density(n1000$cvg.crc), col="#6A3D9A")

```

```{r}
library(ggplot2)
library(dplyr)

make_df <- function(vec, label) {
  data.frame(
    x = vec,
    type = label
  )
}

# Build data for the three n values
all_df <- list()

for (n in c(100, 500, 1000)) {
  
  alpha <- 0.1
  a <- n + 1 - floor((n + 1) * alpha)
  b <- floor((n + 1) * alpha)
  
  # theoretical beta density (sample a grid)
  xgrid <- seq(0.82, 0.98, length.out = 100)
  df_beta <- data.frame(
    x = xgrid,
    density = dbeta(xgrid, a, b),
    type = "Theoretical",
    n = paste0("n_cal = ", n)
  )
  
  # empirical densities
  df_conf <- make_df(get(paste0("n", n))$cvg.conf, "Split")
  df_crc  <- make_df(get(paste0("n", n))$cvg.crc,  "Graphs")
  
  df_conf <- df_conf %>%
    mutate(
      density = density(x, from = 0.82, to = 0.98, n = 100)$y,
      x = density(x, from = 0.82, to = 0.98, n = 100)$x,
      n = paste0("n_cal = ", n)
    )
  
  df_crc <- df_crc %>%
    mutate(
      density = density(x, from = 0.82, to = 0.98, n = 100)$y,
      x = density(x, from = 0.82, to = 0.98, n = 100)$x,
      n = paste0("n_cal = ", n)
    )

  all_df[[as.character(n)]] <- rbind(df_beta, df_conf, df_crc)
}

df_final <- bind_rows(all_df)

# Colors
col_map <- c(
  "Theoretical" = "#7F7F7F",
  "Split" = "#33A02C",
  "Graphs" = "#6A3D9A"
)

df_final$n <- factor(df_final$n,
                     levels = c("n_cal = 100", "n_cal = 500", "n_cal = 1000"))


# Plot
p.density <- ggplot(df_final, aes(x = x, y = density, color = type)) +
  geom_line(size = 0.5) +
  facet_wrap(~ n, nrow = 3, scales = "free_y") +
  scale_color_manual(values = col_map) +
  coord_cartesian(xlim = c(0.82, 0.98)) +
  labs(
    x = "",
    y = "",
    color = ""
  ) +
  theme_minimal(base_size = 15) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank()
  )

```


# Inspect cells with inconclusive predictions
```{r}
spe_test <- spe_other[,-cal]
spe_test$sizes_gr <- as.factor(sizes_gr)
table(spe_test$sizes_gr)
```
```{r}
ggplot(colData(spe_test), aes(sizes_gr, detected))+ geom_violin()
ggplot(colData(spe_test), aes(sizes_gr, sum, fill=sizes_gr))+ geom_violin()+ scale_fill_brewer(palette = "Pastel2")
```

```{r}
ggplot(colData(spe_test)[spe_test$sizes_gr%in% c("1", "15"),], aes(sizes_gr, detected))+ geom_violin()
ggplot(colData(spe_test)[spe_test$sizes_gr%in% c("1", "15"),], 
       aes(sizes_gr, sum, fill=sizes_gr)) + geom_violin()+theme_minimal()+
  ylab("Total counts per cell")+xlab("Size of graph-based set")+
  scale_fill_brewer(palette = "Pastel1") + theme(legend.position = "none")
```


See spatial distribution of inconclusive cells

```{r}
spe_test$inconclusive <- ifelse(spe_test$sizes_gr=="15", "yes", "no")
spe_test$x <- spatialCoords(spe_test)[,1]
spe_test$y <- spatialCoords(spe_test)[,2]
ggplot(data = colData(spe_test), 
       aes(x = x, y = y, color = inconclusive)) +
  geom_point() +
  labs(x = "Spatial Coordinate 1", 
       y = "Spatial Coordinate 2") 
ggplot(data = colData(spe_test), 
       aes(x = x, y = y, color = sizes_gr)) +
  geom_point() +
  labs(x = "Spatial Coordinate 1", 
       y = "Spatial Coordinate 2") 
```

```{r}
# Re-Downloaded to get the images
spe.baysor <- MouseIleumPetukhov2021(segmentation = "baysor")
mem.img <- imgRaster(spe.baysor, image_id = "membrane")
plotXY(colData(spe_test), "inconclusive", mem.img)
```

```{r}
plotXY(colData(spe_test), "cell_type", mem.img)
round(prop.table(table(spe_test$inconclusive, spe_test$cell_type),2),3)
```

```{r}
# See model accuracy stratified by cell type
pr_class <- apply(p_test, 1, function(row) colnames(p_test)[which.max(row)])
mean(pr_class==df_test$Y)

accuracy_by_group <- tapply(seq_along(df_test$Y), df_test$Y, function(idx) {
  mean(pr_class[idx] == df_test$Y[idx])
})

# Print results
accuracy_by_group
round(prop.table(table(spe_test$inconclusive, spe_test$cell_type),2),3)
```


```{r}
spe_rm <- spe.baysor[,spe.baysor$leiden_final %in% c("Removed", "Myenteric Plexus")]
p_rm <- predict(fit, newdata=as.data.frame(t(as.matrix(counts(spe_rm[hvg, ])))),
                type="response")
sets_gr_rm <- getPredictionSets(p_rm, p_cal, df_cal$Y,follow_ontology = TRUE,
                             resample = FALSE, return_sc = FALSE, onto = opi2,
                             labels = unique(df_cal$Y),
                             BPPARAM = MulticoreParam(workers = 4)
                                )
sets_split_rm <- getPredictionSets(p_rm, p_cal, df_cal$Y,follow_ontology = FALSE,
                             resample = FALSE, return_sc = FALSE, onto = opi2,
                             labels = unique(df_cal$Y),
                             BPPARAM = MulticoreParam(workers = 4)
                                )

```

```{r}
sizes <- sapply(sets_gr_rm, length)
round(prop.table(table(sizes, spe_rm$leiden_final),2),3)

sizes_sp <- sapply(sets_split_rm, length)
round(prop.table(table(sizes_sp, spe_rm$leiden_final),2),3)

```


```{r}
pr_rm <- apply(p_rm, 1, function(row) colnames(p_rm)[which.max(row)])
table(pr_rm[spe_rm$leiden_final=="Removed"], sizes[spe_rm$leiden_final=="Removed"])
table(pr_rm[spe_rm$leiden_final=="Myenteric Plexus"], sizes[spe_rm$leiden_final=="Myenteric Plexus"])
```

Common label:
```{r}
spe_test$comm_anc <- vapply(
            sets_graphs,
            function(x) scConform::getCommonAncestor(x, opi2),
            character(1)
        )
table(spe_test$comm_anc)
table(spe_test$sizes_gr)

ch <- rep(NA, length(spe_test$comm_anc))
for(i in seq_along(spe_test$comm_anc))
  ch[i] <- length(scConform:::.children(spe_test$comm_anc[[i]], opi2)) == length(sets_graphs[[i]])
sum(!ch)
```

```{r}
plotXY(colData(spe_test), "comm_anc", mem.img)
```


