---
title: Hierarchical Conformal Prediction with an application to MERFISH mouse ileum
  data
author: "Daniela Corbetta"
date: "2024-06-27"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Code to reproduce analysis in section 4 of the paper "Conformal inference for cell type annotation with graph-structured constraints"

# Load libraries and utility

```{r, message=FALSE}
library(SingleCellExperiment)
library(scater)
library(scran)
library(VGAM)
library(igraph)
library(ontoProc)
library(MerfishData)
library(ExperimentHub)
library(kableExtra)
library(scConform)
`%notin%` = Negate(`%in%`)
```


# Load data and build ontology

Load data and build ontology (for later)

```{r}
# Load data
spe_baysor <- MouseIleumPetukhov2021(
    segmentation = "baysor",
    use.images = FALSE, use.polygons = FALSE
)
# Load ontology
cl <- getOnto("cellOnto", "2023")

tags <- c(
    "CL:0009022", # Stromal
    "CL:0000236", # B cell
    "CL:0009080", # Tuft
    "CL:1000411", # Endothelial
    "CL:1000335", # Enterocyte
    "CL:1000326", # Goblet
    "CL:0002088", # ICC
    "CL:0009007", # Macrophage + DC
    "CL:1000343", # Paneth
    "CL:0000669", # Pericyte
    "CL:1000278", # Smooth Muscle
    "CL:0009017", # Stem + TA
    "CL:0000492", # T (CD4+)
    "CL:0000625", # T (CD8+)
    "CL:0017004" # Telocyte
)
opi <- graph_from_graphnel(onto_plot2(cl, tags))

## Delete instances from other ontologies
sel_ver <- V(opi)$name[c(grep("CARO", V(opi)$name), grep("BFO", V(opi)$name))]
opi1 <- opi - sel_ver

## Rename vertex to match annotations
V(opi1)$name[grep("CL:0000236", V(opi1)$name)] <- "B cell"
V(opi1)$name[grep("CL:1000411", V(opi1)$name)] <- "Endothelial"
V(opi1)$name[grep("CL:1000335", V(opi1)$name)] <- "Enterocyte"
V(opi1)$name[grep("CL:1000326", V(opi1)$name)] <- "Goblet"
V(opi1)$name[grep("CL:0002088", V(opi1)$name)] <- "ICC"
V(opi1)$name[grep("CL:0009007", V(opi1)$name)] <- "Macrophage + DC"
V(opi1)$name[grep("CL:1000343", V(opi1)$name)] <- "Paneth"
V(opi1)$name[grep("CL:0000669", V(opi1)$name)] <- "Pericyte"
V(opi1)$name[grep("CL:1000278", V(opi1)$name)] <- "Smooth Muscle"
V(opi1)$name[grep("CL:0009017", V(opi1)$name)] <- "Stem + TA"
V(opi1)$name[grep("CL:0009022", V(opi1)$name)] <- "Stromal"
V(opi1)$name[grep("CL:0000492", V(opi1)$name)] <- "T (CD4+)"
V(opi1)$name[grep("CL:0000625", V(opi1)$name)] <- "T (CD8+)"
V(opi1)$name[grep("CL:0017004", V(opi1)$name)] <- "Telocyte"
V(opi1)$name[grep("CL:0009080", V(opi1)$name)] <- "Tuft"

## Add the edge from connective tissue cell and telocyte and delete redundant
## nodes
opi1 <- add_edges(opi1, c("connective\ntissue cell\nCL:0002320", "Telocyte"))
gr <- as_graphnel(opi1)
opi2 <- opi1 - c(
    "somatic\ncell\nCL:0002371", "contractile\ncell\nCL:0000183",
    "native\ncell\nCL:0000003"
)

V(opi2)$name <- trimws(gsub("CL:.*|\\n", " ", V(opi2)$name))

gr1 <- as_graphnel(opi2)

## Plot the final ontology
attrs <- list(node = list(shape = "box", fontsize = 15, fixedsize = FALSE))
plot(gr1, attrs = attrs)
```

Preprocess data:
```{r}
spe_baysor$cell_type <- spe_baysor$leiden_final
spe_baysor$cell_type[spe_baysor$cell_type %in% c(
    "B (Follicular, Circulating)",
    "B (Plasma)"
)] <- "B cell"
spe_baysor$cell_type[grep("Enterocyte", spe_baysor$cell_type)] <- "Enterocyte"
spe_baysor <- spe_baysor[, spe_baysor$cell_type %notin% c(
    "Removed",
    "Myenteric Plexus"
)]
spe_baysor

# See frequencies of cell types
table(spe_baysor$cell_type)
```


## Fit a classification model
- Randomly select 500 cells for train dataset
- Fit logit multinomial model on the train dataset
- Split the remaining cells into calibration (1000 cells) and test. See performance of the model on the test dataset

```{r model, warning=FALSE}
# Randomly select 500 cells
set.seed(1703)
train <- sample(1:ncol(spe_baysor), 500)

# Training data
spe_train <- spe_baysor[, train]
spe_train
spe_other <- spe_baysor[, -train]

# get HVGs
spe_train <- logNormCounts(spe_train)
v <- modelGeneVar(spe_train)
hvg <- getTopHVGs(v, n = 50)

# Extract counts and convert data into a data.frame format
df_train <- as.data.frame(t(as.matrix(counts(spe_train[hvg, ]))))
df_other <- as.data.frame(t(as.matrix(counts(spe_other[hvg, ]))))
df_train$Y <- spe_train$cell_type
df_other$Y <- spe_other$cell_type
table(df_train$Y)

# Fit multinomial model
fit <- vglm(Y ~ .,
    family = multinomial(refLevel = "B cell"),
    data = df_train
)

# split data
set.seed(1237)
cal <- sample(1:nrow(df_other), 1000)
df_cal <- df_other[cal,]
df_test <- df_other[-cal,]

# See performance on the test data
p_test <- predict(fit, newdata=df_test, type="response")
pr_class <- apply(p_test, 1, function(row) colnames(p_test)[which.max(row)])
mean(pr_class==df_test$Y)
```

# Compare standard conformal inference with conformal risk control

Obtain prediction sets with standard conformal inference and the graph-based method.

```{r}
p_cal <- predict(fit, newdata=df_cal, type="response")
# Conformal inference
sets_cf <- getPredictionSets(p_test, p_cal, df_cal$Y, follow_ontology = FALSE,
                             resample = FALSE, return_sc = FALSE, 
                             labels = unique(df_cal$Y))
# Check coverage
(cvg_cf <- mean(mapply(function(vec, lst) vec %in% lst, df_test$Y, sets_cf)))

# Graph-based
sets_graphs <- getPredictionSets(p_test, p_cal, df_cal$Y, follow_ontology = TRUE,
                             resample = FALSE, return_sc = FALSE, onto = opi2,
                             labels = unique(df_cal$Y),
                             BPPARAM = MulticoreParam(workers = 6))
# Check coverage
(cvg_graphs <- mean(mapply(function(vec, lst) vec %in% lst, df_test$Y, sets_graphs)))
```


## Comparison of sizes of prediction sets

```{r}
sizes_std <- sapply(sets_cf, length)
sizes_gr <- sapply(sets_graphs, length)
sizes_std_df <- as.data.frame(table(sizes_std))
sizes_gr_df <- as.data.frame(table(sizes_gr))
names(sizes_std_df) <- c("size", "Standard")
names(sizes_gr_df) <- c("size", "Graphs")
merged <- merge(sizes_std_df, sizes_gr_df, by='size', all.x = TRUE, all.y=TRUE)

merged$Graphs[is.na(merged$Graphs)] <- 0
merged$Standard[is.na(merged$Standard)] <- 0

obj <- t(as.matrix(merged[,2:3]))
colnames(obj) <- merged$size

# Fig. 3 in the paper:
barplot(obj, beside=T, legend.text=T,col=c("blue" , "skyblue"))
table(sizes_std)
table(sizes_gr)

mean(sizes_std)
mean(sizes_gr)
```
Bigger prediction sets with graphs

## Comparison of homogeneity of elements inside of prediction sets

```{r}
# Obtain distance matrix
Dshortest <- distances(opi2)
avg.dist.intra.gr <- avg.dist.intra.gr2 <- avg.dist.intra.std <- rep(NA, length(sets_cf))
for(i in 1:length(sets_graphs)){
  dist.gr <- as.matrix(Dshortest[sets_graphs[[i]], sets_graphs[[i]]])
  dist.std <- as.matrix(Dshortest[sets_cf[[i]], sets_cf[[i]]])
  avg.dist.intra.gr[i] <- sum(dist.gr)/(ncol(dist.gr)*(ncol(dist.gr)-1))
  avg.dist.intra.std[i] <- sum(dist.std)/(ncol(dist.std)*(ncol(dist.std)-1))
}

avg.dist.intra.gr[is.na(avg.dist.intra.gr)] <- 0
avg.dist.intra.std[is.na(avg.dist.intra.std)] <- 0

mean(avg.dist.intra.gr, na.rm=T)
mean(avg.dist.intra.std, na.rm=T)

# Lower when working with graphs
```












